\thetitle

### 2013-06-19 tfpie/tfpie.tex
A few edits.

The transition from custom teaching languages to a professional
language takes place about half-way through the course.  At this
point, students have already seen many of the essentials of
object-oriented programming. In particular: object, classes, fields
and methods, dynamic dispatch, inheritance, and overriding.

From this point, almost any language that students might use in future
co-op positions and courses would be an appropriate follow-up.  Our
course transitions to Java, but C\#, Python, Ruby, Eiffel, or
JavaScript would all work naturally.  The key lesson of the transition
is that the fundamental principles underlying object-oriented
programming remain the same between languages, and that learning a new
language is primarily a matter of mapping these concepts to specific
constructs in the new language.  Of course, particular languages also
use unique specific mechanisms which need to be taught to use the
language effectively, but these are rarely as vital as the
cross-language principles.

@@

\begin{enumerate}
\item Explicit representation of unions and interfaces in the
  language.  Previously, interfaces were simply described in stylized
  comments, following the \htdp approach.
\item Types are now specified as part of the program, and are now
  enforced.
\item Java syntax is substantially different and more verbose. For
  example, constructors must be defined explicitly.
\item The testing environment is somewhat different, and requires
  additional boilerplate, although we are
  able to use the JavaLib framework~\cite{local:java-world} to support simple testing by
  structural equality.
\end{enumerate}

\noindent
Of course, there are other differences which cannot be seen from a
code snippet.

@@

Of course, all but the first two of these changes are unrelated to the
fundamental lessons we hope to teach---the rest merely present
additional hurdles for students.

>>>>

The transition from custom teaching languages to a professional
language takes place about half-way through the course.  At this
point, students already have experience with many of the essential
concepts of object-oriented programming. In particular: objects,
classes, fields and methods, dynamic dispatch, inheritance, and
overriding.

From this point, almost any language that students might encounter in
future co-op positions, summer internships, or upper-level courses
would be an appropriate follow-up.  Our course transitions to Java,
but C\#, Python, Ruby, Eiffel, or JavaScript would all work naturally.
The key lesson of the transition is that the fundamental principles
underlying object-oriented programming remain the same between
languages, and that learning a new language is primarily a matter of
mapping these concepts to specific constructs in the new language.  Of
course, particular languages also use unique specific mechanisms which
need to be taught to use the language effectively, but these are
rarely as vital as the cross-language principles.

@@

\begin{enumerate}
\item Explicit representation of unions and interfaces in the
  language.  Previously, interfaces were simply described in stylized
  comments, following the \htdp approach.
\item Types are now specified as part of the program and are (statically)
  enforced.
\item Java syntax is substantially different and more verbose. For
  example, constructors must be defined explicitly.
\item The testing environment is somewhat different, and requires
  additional boilerplate, although we are
  able to use the JavaLib framework~\cite{local:java-world} to support testing with
  structural equality.
\end{enumerate}

\noindent
There are other differences which cannot be seen from a code snippet.

@@

All but the first two of these changes are unrelated to the
fundamental lessons we hope to teach---the rest merely present
additional hurdles for students.
